---
title: "Statistical Analysis of Coworking Space"
author: "Taylor Hughes"
date: "4/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statistical Analyses
First we will run a comparison on 3 sets of data - 

* Proportions of each statistic for Coworking Members
* Proportions of each statistic for Non-Coworking Training data
* Proportions of each statistic for Market Area Data

The outcome of this will allow us to make a hypothesis on:

* Statistics that will have a strong impact on the model
* Which statistical factors would make our Market Area prove to be a weak market for potential customers
* Which statistical factors would make our Market Area prove to be a strong market for potential customers

## Distance to Home
```{r}
table(TrainCensus$CommuteTime, useNA = "ifany")
table(Census$CommuteTime, useNA = "ifany")
table(MarketCensus$CommuteTime, useNA = "ifany")

DTHCommuteTrainYes <- nrow(TrainCensus[TrainCensus$CommuteTime == "Yes"])/nrow(TrainCensus)
DTHTrain <- ggplot(TrainCensus, aes(x = "", y = , fill = CommuteTime)) +
geom_bar(width = 1, stat = "identity")
DTHTrain
```

